---
title: "LFQ comparison"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(dplyr)
library(stringr)
library(limma)
library(PECA)
library(vsn)
library(pROC)
library(kableExtra)
library(ggplot2)
```

## ABRF

Let's define which proteins to examine:

```{r}
# ABRF-1: Beta Galactosidase from E Coli
# ABRF-2: Lysozyme from Gallus gallus
# ABRF-3: Amylase from Aspergillus
# ABRF-4: Protein G Streptococcus
abrfprots <- c("BGAL", "LYSC", "ASPNC", "PROTEING")
```


## FlashLFQ
### Without match-between-runs (no MBR and no normalization)
```{r}
flash <- read.delim("data/Quant_Paper_dec_results/ABRF/FLASHLFQ/noMBR\ without\ normalization/Galaxy284-[FlashLFQ_on_data_99,_data_53,_and_others__QuantifiedPeptides.tsv].tabular", stringsAsFactors = FALSE) %>%
                mutate(Sequence = gsub("<cmm>", "", str_match(Sequence, "NH2-(.+)-COOH")[,2]))

# Aggregate intensities for each protein and each sample
flash_table <- t(sapply(abrfprots, function(x){ 
    flash %>% filter(grepl(x,Protein.Groups)) %>%
    group_by(Protein.Groups) %>%
    summarise_if(is.numeric, sum)
  }))

flash_table <- flash_table[,c(3,4,2,5)]
flash_table
write.table(flash_table, file = "results/ABRF/FlashLFQ/intensities_flash.tsv", quote=FALSE, sep="\t")
```

### With match-between-runs (with MBR and no normalization)
```{r}
flash_mbr <- read.delim("data/Quant_Paper_dec_results/ABRF/FLASHLFQ/MBR\ without\ normalization/Galaxy272-[FlashLFQ_on_data_99,_data_53,_and_others__QuantifiedPeptides.tsv].tabular", stringsAsFactors = FALSE) %>%
  mutate('Full Sequence' = gsub("NH2-", "",
                         gsub("-COOH", "",
                              gsub("<cmm>","",
                                   gsub("<ox>","", 'Full Sequence')))))
flash_mbr
# Aggregate intensities for each protein and each sample
flash_mbr_table <- t(sapply(abrfprots, function(x){ 
    flash_mbr %>% filter(grepl(x,Protein.Groups)) %>%
    group_by(Protein.Groups) %>%
    summarise_if(is.numeric, sum)
  }))

flash_mbr_table <- flash_mbr_table[,c(3,4,2,5)]
flash_mbr_table
write.table(flash_mbr_table, file = "results/ABRF/FlashLFQ/intensities_flash_mbr.tsv", quote=FALSE, sep="\t")
```

## FlashLFQ With normalization
### Without match-between-runs (no MBR with normalization)
```{r}
flash_norm <- read.delim("data/Quant_Paper_dec_results/ABRF/FLASHLFQ/noMBRwith\ normalization/Galaxy278-[FlashLFQ_on_data_99,_data_53,_and_others__QuantifiedPeptides.tsv].tabular", stringsAsFactors = FALSE) %>%
                mutate(Sequence = gsub("<cmm>", "", str_match(Sequence, "NH2-(.+)-COOH")[,2]))

# Aggregate intensities for each protein and each sample
flash_norm_table <- t(sapply(abrfprots, function(x){ 
    flash %>% filter(grepl(x,Protein.Groups)) %>%
    group_by(Protein.Groups) %>%
    summarise_if(is.numeric, sum)
  }))

flash_norm_table <- flash_norm_table[,c(3,4,2,5)]
flash_norm_table
write.table(flash_norm_table, file = "results/ABRF/FlashLFQ/intensities_flash_norm.tsv", quote=FALSE, sep="\t")
```

### With match-between-runs (with MBR, with normalization)
```{r}
flash_mbr_norm <- read.delim("data/Quant_Paper_dec_results/ABRF/FLASHLFQ/MBR\ with\ normalization/Galaxy266-[FlashLFQ_on_data_99,_data_53,_and_others__QuantifiedPeptides.tsv].tabular", stringsAsFactors = FALSE) %>%
  mutate('Full Sequence' = gsub("NH2-", "",
                         gsub("-COOH", "",
                              gsub("<cmm>","",
                                   gsub("<ox>","", 'Full Sequence')))))

# Aggregate intensities for each protein and each sample
flash_mbr_norm_table <- t(sapply(abrfprots, function(x){ 
    flash_mbr %>% filter(grepl(x,Protein.Groups)) %>%
    group_by(Protein.Groups) %>%
    summarise_if(is.numeric, sum)
  }))

flash_mbr_norm_table <- flash_mbr_norm_table[,c(3,4,2,5)]
flash_mbr_norm_table
write.table(flash_mbr_norm_table, file = "results/ABRF/FlashLFQ/intensities_flash_norm_mbr.tsv", quote=FALSE, sep="\t")
```


## moFF
### Without match-between-runs (no MBR)

```{r}
# When moFF is run without MBR, we need to combine all of the individual peptide summaries, which we do with `join_all` from `plyr`. 
moff_files <- list.files('data/Quant_Paper_dec_results/ABRF/moFF/noMBR\ without\ normalization/peptide_summary/',
                    full.names = TRUE)
moff_no_mbr <- lapply(moff_files, function(i) read.delim(i, as.is = TRUE) %>%
                rename(!!paste("Intensity_",
                     str_extract(i, "S[0-9_]"), sep="") := sumIntensity_Peptide)) %>%
                join_all(by=c("peptide", "prot"), type = "full") %>%
                rename(Sequence = peptide)

# Aggregate intensities for each protein and each sample
moff_no_mbr_table <- t(sapply(abrfprots, function(x){ 
    output <- moff_no_mbr %>% filter(grepl(x, prot)) %>%
      group_by(prot)
    output[is.na(output)] <- 0
    output %>% summarise_if(is.double, sum)
}))

moff_no_mbr_table <- moff_no_mbr_table[,c(2,3,5,4)]
moff_no_mbr_table
write.table(moff_no_mbr_table, file = "results/ABRF/moFF/intensities_moff_no_mbr.tsv", quote=FALSE, sep="\t")
```

### With match-between-runs (with MBR)
```{r}
moff_mbr <- read.delim("data/Quant_Paper_dec_results/ABRF/moFF/MBR\ without\ normalization/Galaxy119-[moFF_on_data_115,_data_114,_and_others__peptide_summary].tabular", stringsAsFactors = FALSE)
moff_mbr
# Aggregate intensities for each protein and each sample
moff_mbr_table <- t(sapply(abrfprots, function(x){ 
    moff_mbr %>% filter(grepl(x, prot)) %>%
    group_by(prot) %>%
    summarise_if(is.numeric, sum)
  }))
# moff_mbr_table

moff_mbr_table <- moff_mbr_table[,c(3,5,2,4)]
moff_mbr_table
write.table(moff_mbr_table, file = "results/ABRF/moFF/intensities_moff_mbr.tsv", quote=FALSE, sep="\t")
```

## MaxQuant
### Without match-between-runs (no MBR) 
```{r}
maxquant <- read.delim("data/Quant\ Paper\ Redone_11102019/ABRF/MaxQuant/ABRF_MQ_noMBR/combined/txt/peptides.txt", stringsAsFactors = FALSE)

# Aggregate intensities for each protein and each sample
maxquant_table <- t(sapply(abrfprots, function(x){ 
  if (x == "BGAL"){
    x <- "BGAL_ECOLI"
  }
  maxquant %>% filter(grepl(x, Proteins)) %>%
    select(Proteins, Intensity.1, Intensity.2, Intensity.3, Intensity.4) %>%
    group_by(Proteins) %>%
    summarise_if(is.numeric, sum)
}))

maxquant_table <- maxquant_table[,c(3,4,2,5)]
maxquant_table
write.table(maxquant_table, file = "results/ABRF/MaxQuant/intensities_maxquant.tsv", quote=FALSE, sep="\t")
```

### With match-between-runs (with MBR)
```{r}
maxquant_mbr <- read.delim("data/Quant\ Paper\ Redone_11102019/ABRF/MaxQuant/ABRF_MQ_withMBR/combined/txt/peptides.txt", stringsAsFactors = FALSE)

# Aggregate intensities for each protein and each sample
maxquant_mbr_table <- t(sapply(abrfprots, function(x){ 
  if (x == "BGAL"){
    x <- "BGAL_ECOLI"
  }
  maxquant_mbr %>% filter(grepl(x, Proteins)) %>%
    select(Proteins, Intensity.1, Intensity.2, Intensity.3, Intensity.4) %>%
    group_by(Proteins) %>%
    summarise_if(is.numeric, sum)
}))

maxquant_mbr_table <- maxquant_mbr_table[,c(3,4,2,5)]
maxquant_mbr_table
write.table(maxquant_mbr_table, file = "results/ABRF/MaxQuant/intensities_maxquant_mbr.tsv", quote=FALSE, sep="\t")
```
