---
title: "LFQ comparison"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(dplyr)
library(stringr)
library(limma)
library(PECA)
library(vsn)
library(pROC)
library(kableExtra)
library(ggplot2)
```

## True Ratios

The 48 UPS proteins (Sigma Aldrich) were spiked into an *E. coli* background. UPS1 has all 48 UPS proteins at the same concentration, while UPS2 has varying concentrations. We define the ratio as UPS2 / UPS1, so the true ratios run from 10 to $10^{-4}$ -- in addition, 8 proteins have the same concentration in UPS2 and UPS1. The proteins and their associated UPS2/UPS1 ratios are defined below:

```{r}
ups_protein_ids <- c(
    #ratio of 10
    "P00915", "P00918", "P01031", "P69905", "P68871", "P41159", "P02768", "P62988",
    # ratio of 1
    "P04040", "P00167", "P01133", "P02144", "P15559", "P62937", "Q06830", "P63165",
    # ratio of 0.1
    "P00709", "P06732", "P12081", "P61626", "Q15843", "P02753", "P16083", "P63279",
    # ratio of 0.01
    "P01008", "P61769", "P55957", "O76070","P08263","P01344","P01127","P10599",
    # ratio of 0.001
    "P99999","P06396","P09211","P01112","P01579","P02787","O00762","P51965",
    # ratio of 0.0001
    "P08758","P02741","P05413","P10145","P02788","P10636","P00441","P01375"
)

ups_true_df <- data.frame(prot = ups_protein_ids,
                          ratios = rep(
                            sapply(1:-4, function(factor){
                              return(10^factor)
                            }),
                            each = 8),
                          stringsAsFactors = FALSE)

# find the protein ids for proteins that are constant
ups_non_de <- filter(ups_true_df, ratios == 1)$prot
```

## Quantification Results

We analyze 6 different result files on the peptide level, and 2 different result files on the protein level. The peptide files are as follows:

* moFF's peptide summary, produced from MaxQuant's `msms.txt` - run with/without MBR (2 files)
* FlashLFQ's peptide file: `msms_FlashLFQ_QuantifiedBaseSequences.tsv` - produced from MaxQuant's `msms.txt` and run with/without MBR (2 files)
* MaxQuant's peptide file: `peptides.txt`, from MaxQuant, with/without MBR (2 files)

The 2 protein files are MaxQuant's protein report (`proteinGroups.txt`), from MaxQuant run with/without MBR. 

In the following, we read in all of the results files. In most cases, the protein IDs are in a list, separated by ";", so we use the function `get_protein()` to get the first protein in the list. 
```{r}
get_protein <- function(prot_vec){ sapply(prot_vec, function ( row ) { return(str_extract(row, "[A-Z0-9]{6}")) }) }
```

We don't want any character columns to be read as factors:
```{r}
options(stringsAsFactors = FALSE)
```

Let's also set up a function for getting the filename using a regular expression:
```{r}
root <- "Jan_Quant_Paper_revised"
get_filename <- function(root, directory, regex){
  directory <- paste0("data/", root, directory)
  filename <- paste0(directory, list.files(directory, pattern=regex))
  print(filename)
  return(filename)
}
```

### FlashLFQ (no normalization)
```{r flash}
flash <- read.delim(get_filename(root, "/UPS/FLASHLFQ/MBR_without_norm/", ".+QuantifiedPeptides.tsv].tabular"),
                    stringsAsFactors = FALSE) %>%
    rename(peptide = Sequence, prot = Protein.Groups) %>%
    select(peptide, prot, starts_with("Intensity"))
flash$prot <- get_protein(flash$prot)
flash_names <- names(flash)
flash$peptide <- gsub("<cmm>","",
                      gsub("<ox>","",
                           gsub("NH2-","", gsub("-COOH","", flash$peptide))))
#flash

# flash no mbr
flash_no_mbr <- read.delim(get_filename(root, "/UPS/FLASHLFQ/noMBR_without_norm/", ".+QuantifiedPeptides.tsv].tabular"),
                           stringsAsFactors = FALSE) %>%
    rename(peptide = Sequence, prot = Protein.Groups) %>%
    select(peptide, prot, starts_with("Intensity"))
flash_no_mbr$prot <- get_protein(flash_no_mbr$prot)
flash_no_mbr$peptide <- gsub("<cmm>","",
                      gsub("<ox>","",
                           gsub("NH2-","", gsub("-COOH","", flash_no_mbr$peptide))))
#flash_no_mbr
```

### FlashLFQ (with normalization)
```{r flash_norm}
flash_norm <- read.delim(get_filename(root, "/UPS/FLASHLFQ/MBR_with_norm/",
                                      ".+BayesianFoldChangeAnalysis.tsv].tabular"),
                    stringsAsFactors = FALSE) %>%
    rename(prot = Protein.Group)# %>%
    #select(prot, Protein.Intensity.for.Treatment.Condition)
flash_norm$prot <- get_protein(flash_norm$prot)
#flash_names <- names(flash)
#flash_norm

# flash no mbr
flash_no_mbr_norm <- read.delim(get_filename(root, "/UPS/FLASHLFQ/noMBR_with_norm/",
                                             ".+BayesianFoldChangeAnalysis.tsv].tabular"),
                            stringsAsFactors = FALSE) %>%
    rename(prot = Protein.Group)
flash_no_mbr_norm$prot <- get_protein(flash_no_mbr_norm$prot)
#flash_no_mbr_norm
```


### moFF
```{r moff}
moff <- read.delim(get_filename(root, "/UPS/moff/withMBR/", ".+peptide_summary].tabular"),
                   stringsAsFactors = FALSE)
# replace protein list with first protein
moff$prot <- get_protein(moff$prot)
moff_names <- names(moff)
moff <- moff[,order(colnames(moff))]
#moff
```

When moFF is run without MBR, we need to combine all of the individual peptide summaries, which we do with `join_all` from `plyr`. 

```{r moff_no_mbr}
# moff, no mbr
moff_files <- list.files(paste0('data/',root, '/UPS/moff/withoutMBR/peptide_summary/'),
                    full.names = TRUE)
#tail(unlist(str_split(moff_files[1], "/")), n=1)

moff_no_mbr <- lapply(moff_files, function(i) df = read.delim(i, as.is = TRUE) %>% 
  rename(!!paste("sumIntensity_20130510_EXQ1_IgPa_QC_",
                 str_extract(i, "UPS[0-9_]+"), sep="") := sumIntensity_UPS_PSM_All_edited)) %>%
                        join_all(by=c("peptide", "prot"), type = "full")
moff_no_mbr$prot <- get_protein(moff_no_mbr$prot)
#moff_no_mbr
```

### MaxQuant

#### Peptides
```{r mqpep}
mq <- read.delim(paste0('data/', root, "/UPS/MAXQUANT/MQ_MBR/txt/peptides.txt"),
                 stringsAsFactors = FALSE) %>%
    rename(peptide = Sequence, prot = Proteins) %>%
    select(peptide, prot, starts_with("Intensity."))
mq$prot <- get_protein(mq$prot)
mq_names <- names(mq)
#mq
# mq no mbr
mq_no_mbr <- read.delim(paste0('data/', root, "/UPS/MAXQUANT/MQ_NoMBR/txt/peptides.txt"),
                        stringsAsFactors = FALSE) %>%
    rename(peptide = Sequence, prot = Proteins) %>%
    select(peptide, prot, starts_with("Intensity."))
mq_no_mbr$prot <- get_protein(mq_no_mbr$prot)
#mq_no_mbr
```

#### Proteins

With the MaxQuant protein reports, we are going to do a t-test on each row of the data matrix, rather than using `PECA` as for the peptide-level analysis. Therefore, we filter the results to keep only the proteins that had values in 3 out of 4 replicates for both groups, which is the minimum necessary for a robust T-test. 

```{r mqprot}
mq_prot_analysis <- function(proteinGroups, name){
    #don't remove contaminants, because some UPS proteins are potential contam
    mq_prot <- read.delim(proteinGroups,
                      na.strings = c("0", "", "NA"),
                      stringsAsFactors = FALSE) %>%
        filter(is.na(Reverse)) %>% # if not decoy, is na
        mutate(prot = get_protein(Protein.IDs),
               ups_prot_id = str_split(prot, "ups\\|", simplify = TRUE)[, 1]) %>%
        mutate(ups = prot %in% ups_protein_ids, de = ups & !(prot %in% ups_non_de)) %>%
        select(prot, ups, de, starts_with("LFQ"), ups_prot_id)
    
    # filter to 3 or more observations
    mq_prot$enough_obs <- rowSums(is.na(mq_prot[4:7])) <= 1 &
        rowSums(is.na(mq_prot[8:11])) <= 1
    mq_prot_filt <- filter(mq_prot, enough_obs) %>% select(-enough_obs)
    
    # t testing
    mq_ts <- rep(0, nrow(mq_prot_filt))
    for (i in 1:nrow(mq_prot_filt)){
        mq_ts[i] <- t.test(x = mq_prot_filt[i, 4:7], y = mq_prot_filt[i, 8:11])$p.value
    }
    mq_prot_filt$p.fdr <- p.adjust(mq_ts, method = "fdr")
    mq_prot_filt$sr <- rowMeans(mq_prot_filt[, 8:11], na.rm = TRUE)/
        rowMeans(mq_prot_filt[, 4:7], na.rm = TRUE)
    # mq_roc <- roc(mq_prot_filt$de, mq_prot_filt$p.fdr)
    joined_mq <- filter(mq_prot_filt, ups) %>%
        left_join(ups_true_df, by = c("ups_prot_id" = "prot")) 
    rmsle <- sqrt(mean((log10(joined_mq$sr) - log10(joined_mq$ratios))^2, na.rm = TRUE))
    mq_data <- data.frame("norm_methods" = "MaxLFQ", "quant_methods" = name, rmsle = rmsle)
    list("full_df" = mq_prot_filt, "rmlse" = mq_data, "joined_mq" = joined_mq)
}

# mq prot
mq_data <- mq_prot_analysis(paste0('data/', root, "/UPS/MAXQUANT/MQ_MBR/txt/proteinGroups.txt"), "MaxQuant")

# mq prot, no mbr
mq_no_mbr_data <- mq_prot_analysis(paste0('data/', root, "/UPS/MAXQUANT/MQ_NoMBR/txt/proteinGroups.txt"), "MaxQuant_no_MBR")
```

## Differential expression analysis and ratio estimation

Below is the function that is used on all of the peptide reports to normalize and then test for differential expression. The R/Bioconductor package `limma` is used for normalization, and the package `PECA` is used to "roll-up" from peptides to proteins and to test for differential expression (using a modified t-test). Then, `roc` from the `pROC` package is used to create a ROC curve, from which the AUC is estimated. 

```{r testfun}
peptides_normalize_and_test_de <- function(df, quant_method, int_col_vec, id_col_name,
                                           norm_method, grp1_col_name, grp2_col_name){
  if (str_detect(quant_method, "norm")){
    df_peca <- df %>% rename(p = False.Discovery.Rate,
                             slr = Protein.Log2.Fold.Change)
  } else {
    if (norm_method == "vsn"){
      intensities <- 2^limma::normalizeVSN(as.matrix(df[, int_col_vec]))
    } else if (norm_method != 'NA') {
      intensities <- 2^(limma::normalizeBetweenArrays(log2(as.matrix(df[, int_col_vec])), method = norm_method))
    } else {
      intensities <- as.matrix(df[, int_col_vec])
    }
    
    df_norm <- data.frame(prot = df$prot, intensities, stringsAsFactors = FALSE)
    df_peca <- PECA::PECA_df(df_norm, id = id_col_name, samplenames1 = grp1_col_name, samplenames2 = grp2_col_name,
                             test = "modt")
    df_peca$prot <- rownames(df_peca)
  }
  
  # ups proteins
  df_peca$ups <- as.numeric(df_peca$prot %in% ups_protein_ids)
  # ups proteins that are not DE
  for (i in 1:nrow(df_peca)){
    df_peca$de[i] <- ifelse(df_peca$prot[i] %in% ups_non_de, 0, df_peca$ups[i])
  }
  
  df_peca_ratio <- df_peca %>%
    filter(ups == 1) %>%
    left_join(ups_true_df, by = "prot") %>%
    mutate(sr = 2^slr) # peca returns log2 fold change
  sqr_err <- (log10(df_peca_ratio$sr) - log10(df_peca_ratio$ratios))^2
  rmsle <- sqrt(mean(sqr_err, na.rm = TRUE))
  
  return(list("peca" = df_peca_ratio,
              "full_peca" = df_peca,
              "rmsle" = rmsle))
}
```
```{r peptides, warning=FALSE,message=FALSE}
norm_methods <- c("cyclicloess", "scale", "quantile", "vsn")
quant_methods <- c("moFF",  "moFF_no_MBR", "FlashLFQ",
                   "FlashLFQ_no_MBR", "MaxQuant", "MaxQuant_no_MBR")
quants <- list("moFF" = moff,  "moFF_no_MBR" = moff_no_mbr,
               "FlashLFQ" = flash, "FlashLFQ_no_MBR" = flash_no_mbr,
               "MaxQuant" = mq, "MaxQuant_no_MBR" = mq_no_mbr,
               "FlashLFQ_norm" = flash_norm, "FlashLFQ_no_MBR_norm" = flash_no_mbr_norm)
method_list <- expand.grid(list(norm_methods = norm_methods, quant_methods = quant_methods), stringsAsFactors = FALSE)
method_list <- rbind(method_list,
                     c('NA', 'FlashLFQ_norm'),
                     c('NA', 'FlashLFQ_no_MBR_norm'))
results <- vector(length = nrow(method_list), mode = "list")
# aucs <- rep(0, nrow(method_list))
rmses <- rep(0, nrow(method_list))
for (i in 1:nrow(method_list)){
   if (i != 14){
    df <- quants[[method_list$quant_methods[i]]]
    names_df <- names(df)
    ith_result <- peptides_normalize_and_test_de(df,
                                                 quant_method = method_list$quant_methods[i],
                                                 int_col_vec = 3:10,
                                                 id_col_name = 'prot',
                                                 norm_method =  method_list$norm_methods[i],
                                                 grp1_col_name = names_df[7:10],
                                                 grp2_col_name = names_df[3:6])
    results[[i]] <- ith_result
    # aucs[i] <- ith_result$roc$auc
    rmses[i] <- ith_result$rmsle
   }
}
all_df <- cbind(method_list, rmsle = rmses)
```

# Results

## RMSLE

### Combined Levels
```{r proteins}
fc_accuracy <- rbind(all_df, mq_data$rmlse, mq_no_mbr_data$rmlse) %>%
    rename("Normalization" = "norm_methods",
           "Tool" = "quant_methods") 
fc_accuracy
write.table(fc_accuracy, file = "results/fold_change_accuracy.tabular",
            quote = FALSE,
            row.names = FALSE,
            sep = '\t')
```

### Individual ratio levels
```{r}
all_himedlo <- data.frame(t(sapply(c(1:(length(results) + 1)), function(i){
  if (i >= length(results)){
    norm_method <- "MaxLFQ"
    if (i == length(results)){
      quant_method <- "MaxQuant"
      peca_loc <- mq_data$joined_mq
    } else {
      quant_method <- "MaxQuant_no_MBR"
      peca_loc <- mq_no_mbr_data$joined_mq
    }
  } else {
    if (i > 13){ i <- i + 1 }
    norm_method <- method_list$norm_methods[i]
    quant_method <- method_list$quant_methods[i]
    peca_loc <- results[[i]]$peca
  }
  results <- sapply(c(1:-4), function(exp){
    de_level <- peca_loc %>% filter(ratios == 10^exp)
    num <- nrow(filter(de_level, !is.na(sr)))
    rsmle <- sqrt(mean((log10(de_level$ratios) - log10(de_level$sr))^2, na.rm = TRUE))
    return(c(rsmle,num))
  })
  return(c(norm_method, quant_method, results))
})))
headers <- unlist(sapply(1:-4, function(exp){
  ratio <- as.character(10^exp)
  return(list(ratio, paste0(ratio, 'n')))
}))
colnames(all_himedlo) <- c('norm_methods', 'quant_methods', headers)
all_himedlo
write.table(all_himedlo, file="results/high_low_fold_change_accuracy_new.tab",
            sep="\t", quote=FALSE, row.names=FALSE)
``` 

